### 2018-08-19 paper/preliminaries.tex
checked the fixes

As shown in~\cite{Roscoe:1994:CME:197600}, any finite-state CSP process $P$
can be represented by a \emph{normalised transition graph}
$$
G(P) = ( N, \ii n, \Sigma, t : N\times\Sigma \pfun N, r : N \fun \mathbb{P}\mathbb{P}(\Sigma)),
$$
with nodes $N$, initial node $\ii n\in N$, and process alphabet $\Sigma$. The
partial \emph{transition function} $t$ maps a node $n$ and an event
$e\in\Sigma$ to its successor node $t(n,e)$. If $(n,e)$ is in the domain of
$t$, then there is a transition from $n$ with label $e$. Normalisation of
$G(P)$ is reflected by the fact that $t$ is a function. The graph construction
described in~\cite{Roscoe:1994:CME:197600} implies that all nodes $n$ in $N$ are reachable 
by a sequence of edges labelled by $e_1\dots e_k$ and connecting states 
$\ii n,n_1,\dots,n_{k-1},n$, such that
\[
n_1 = t(\ii n,e_1), \quad n_i = t(n_{i-1},e_i),\ i = 2,\dots,k-1,\quad
n= t(n_{k-1},e_k).
\]

A finite sequence of events $s\in\Sigma^*$ is a \emph{trace} of $P$, if and only if 
there
is a path through $G(P)$ starting  at $\ii n$ whose edge labels coincide with
$s$. The set of traces of $P$ is denoted by $\trc(P)$. We use the analgous notation
$\trc(G(P))$ for the set of finite, initialised paths through $G(P)$, each path represented by its finite sequence of edge labels.
If $s\in\trc(P)$, then
the process corresponding to $P$ after having executed $s$ is denoted by
$P/s$. Since $G(P)$ is normalised, there is a unique node reached by applying
the events from $s$ one by one, starting in $\ii n$. Therefore, $G(P)/s$  is
also well defined. By $[n]^0$ we denote the \emph{initials} of $n$:~the set
of events occurring as labels in any outgoing transitions.
$$
[n]^0 = \{ e\in\Sigma~|~(n,e)\in\dom~t \}
$$
We also use this notation for CSP processes:~$[P]^0$ is the set of events $P$
may engage into, in other words, the initials of $P$ after the empty trace of
events, that is, $initials(P/\langle\rangle)$ as defined
in~\cite{Roscoe2010}, for example.

@@

When investigating  tests for failures refinement, the notion of
\emph{acceptances}~\cite{Hennessy:1988:ATP:50497}, which is dual to refusals,
is also useful. An acceptance set of $P/s$ is a subset of the initials
$[P/s]^0$, i.e., a subset of events labelling  outgoing transitions of
$G(P)/s$. If the behaviour of  $P/s$ is deterministic, its only acceptance
equals $[P/s]^0$, because $P/s$ never refuses any of the events contained in
this set. If $P/s$ is nondeterministic, it internally chooses one of its
\emph{minimal acceptance sets} $A$ and never refuses any event in $A$, while
refusing the events in $\Sigma-A$. The acceptances of $P/s$ are denoted by
$\accs(P/s)$, and the minimal acceptances by $\minaccs(P/s)$. They satisfy
the following properties.
%
\begin{eqnarray}
A\in \minaccs(P/s) & \Leftrightarrow & \exists R\in\maxrefs(P/s) \wedge A = \Sigma-R
\label{eq:accref}
\\
\bigcup \{ A~|~A\in \accs(P/s)\} & = & [P/s]^0
\label{eq:accinitials}
\\
 X\in\accs(P/s) &\Leftrightarrow & A\in \minaccs(P/s) \wedge A\subseteq X \subseteq [P/s]^0
 \label{eq:accsubset}
\end{eqnarray}
%
Note that (\ref{eq:accref}) can be regarded as a definition of minimal
acceptances by means of maximal refusals. In this case,
(\ref{eq:accinitials}) and (\ref{eq:accsubset}) are consequences of this
definition and the fact that refusals are subset-closed. \fixme{alcc: are you
sure? (4) does not talk about Acc. Is this important?}
\fixme{jp: I'm sure and I can prove this. However, we don't need this fact, so I suggest to
removed it.}

Exploiting formulas (\ref{eq:accref},\ref{eq:accinitials},\ref{eq:accsubset}),
every node of a normalised transition graph can alternatively be labelled
with their minimal acceptances, and this information is equivalent to that
contained in the maximal refusals. Since process states $P/s$ are equivalently expressed
by states $G(P)/s$ of $P$'s normalised transition graph, we also write
$\minaccs(G(P)/s)$ and $\maxrefs(G(P)/s)$ to denote the minimal acceptences and maximal refusals, respectively, labelling graph state $G(P)/s$ and note 
that $\minaccs(G(P)/s) = \minaccs(P/s)$ and $\maxrefs(G(P)/s) = \maxrefs(P/s)$.

@@

Refinement relations between finite-state CSP processes $P, Q$ can be be
expressed by means of their normalised transition graphs in the following
way. \fixme{alcc: trc was defined for P, not G(P) and this lemma is used
later for trc(P). So, we need to be clearer as to what we mean.}
\fxnote{jp: I have now introduced notation $\trc(G(P))$ above, so I think that the
statement of the lemma is now sufficiently clear.}
%
\begin{lemma}
  \label{lemma:tgtrcref}
  \begin{eqnarray}
  P \lessdet_T Q & \Leftrightarrow & \trc(G(Q)) \subseteq\trc(G(P))
  \label{eq:trcrefa}
  \\
  \label{eq:failrefa}
  P \lessdet_F Q & \Leftrightarrow & \trc(G(Q)) \subseteq\trc(G(P)) \wedge {} \nonumber
  \\ & & \forall s\in\trc(G(Q)), R_Q\in\maxrefs(G(Q)/s):  \nonumber
  \\ & & \tabd
  \exists R_P\in\maxrefs(G(P)/s): R_Q\subseteq R_P
  \\
  \label{eq:failrefb}
  P \lessdet_F Q & \Leftrightarrow & \trc(G(Q)) \subseteq\trc(G(P)) \wedge {} \nonumber
  \\ & & \forall s\in\trc(G(Q)), A_Q\in\minaccs(G(Q)/s): \nonumber
  \\ & & \tabd
  \exists A_P\in\minaccs(G(P)/s): A_P\subseteq A_Q
  \end{eqnarray}
  \xbox
\end{lemma}
%

Given a finite-state CSP process $P$ and its normalised transition graph
\[
G(P) = ( N, \ii n, \Sigma, t : N\times\Sigma \pfun N, r : N \fun \mathbb{P}\mathbb{P}(\Sigma)),
\]
suppose that $V\subseteq\Sigma^*$ is a
prefix-closed set  of sequences of events. By $t(\ii n,V)$ we denote the set
\[
t(\ii n,V) = \{ n\in N~|~\exists s\in V: s\in\trc(P)\wedge G(P)/s = n \}
\]
of nodes in $N$ that are reachable in $G(P)$ by applying traces of $V$.
The following lemma specifies a simple construction method for such sets $V$ reaching 
{\it every} node of $N$.

\begin{lemma}
\label{lemma:extendV} Let $P$ be a CSP process with normalised transition
graph $G(P)$. \fixme{alcc: can a
normalised graph for a process have states that are unreachable?}
\fxnote{jp: You are rights: all states in G(P) are reachable by construction of the graph -- I have added this information to the section introducing the graphs.}
Let
$V\subseteq\Sigma^*$ be a finite prefix-closed set of sequences of events.
Suppose that  $G(P)$ reaches $k < |N|$ nodes under $V$, that is, $|t(\ii
n,V)| = k$. Let $V.\Sigma$ denote the set of all sequences from $V$, extended
by any event of $\Sigma$. Then $G(P)$ reaches at least $(k+1)$ nodes under
$V\cup V.\Sigma$.
\end{lemma}
\begin{proof}
Suppose that $n'\in (N - t(\ii n,V))$.  Since all nodes in $N$ are reachable,
there exists a trace $s$ such that $G(P)/s = n'$. Decompose $s = s_1.e.s_2$
with $s_i\in\Sigma^*, e\in\Sigma$, such that $G(P)/s_1 \in t(\ii n,V)$ and
$G(P)/s_1.e \not\in t(\ii n,V)$. Such a decomposition always exists, because
$V$ is prefix-closed and therefore contains the empty trace $\varepsilon$.
Note, however, that it is not necessarily the case that $s_1\in V$.

Since $G(P)$ reaches $G(P)/s_1$ under $V$, there exists a trace $u\in V$ such
that $G(P)/u = G(P)/s_1 = \ol n$. Since $s = s_1.e.s_2$ is a trace of $P$ and
$G(P)/s_1 = \ol n$, then $(\ol n,e)$ is in the domain of $t$. So, $ G(P)/u.e
= G(P)/s_1.e = n$ is a well-defined node of $N$ not contained in $t(\ii
n,V)$. Since $u.e\in V\cup V.\Sigma$, $G(P)$ reaches at least the additional
node $n$ under $V\cup V.\Sigma$. This completes the proof. \xbox
\end{proof}
\fixme{alcc: Explain the structure of the section? Can the above result be in
a section, rather than the introduction of the section?}
\fxnote{jp: I have moved this lemma now to the section on preliminaries, this should make the main section on the finite complete test suite easier to read.}

@@

To prove ``$2 \Rightarrow 1$'', assume $P\lessdet_T Q$, but $P\not\lessdet_F
Q$. According to Lemma~\ref{lemma:tgtrcref}, (\ref{eq:failrefb}), there
exists $s\in\trc(P)\cap \trc(Q)$ such that
\[
\exists A_Q\in\minaccs(G(Q)/s): \forall A_P\in\minaccs(G(P)/s): A_P\not\subseteq A_Q
\qquad (*)
\]
Let $A$ be such a set $A_Q$ fulfilling (*).
Define
\[
\overline H = \bigcup\{ A_P \setminus A~|~A_P\in\minaccs(G(P)/s) \}.
\]
Since $A_P \setminus A \neq\varnothing$ for all $A_P$ because of (*),
$\overline H$ is a hitting set of $\minaccs(G(P)/s)$ which has an  empty
intersection with $A_Q$. \fixme{alcc: minaccs is used for P above.}
\fxnote{jp: Above Example 1, I have now introduced the notation $\minaccs(G(P)/s)$ and
$\maxrefs(G(P)/s)$.}
Minimising $\overline H$ induces the existence of a minimal hitting set $H\in
\text{minHit}(P/s)$ which is {\it not} a hitting set of $\minaccs(G(Q)/s)$, a
contradiction to Assumption~2. This completes the proof of the lemma. \xbox
\end{proof}
%
This result is used in the sequel in Section~\ref{sec:finitecompletefails}.

>>>>

As shown in~\cite{Roscoe:1994:CME:197600}, any finite-state CSP process $P$
can be represented by a \emph{normalised transition graph}
$$
G(P) = ( N, \ii n, \Sigma, t : N\times\Sigma \pfun N, r : N \fun \mathbb{P}\mathbb{P}(\Sigma)),
$$
with nodes $N$, initial node $\ii n\in N$, and process alphabet $\Sigma$. The
partial \emph{transition function} $t$ maps a node $n$ and an event
$e\in\Sigma$ to its successor node $t(n,e)$. If $(n,e)$ is in the domain of
$t$, then there is a transition from $n$ with label $e$. Normalisation of
$G(P)$ is reflected by the fact that $t$ is a function. The graph construction
in~\cite{Roscoe:1994:CME:197600} implies that all nodes $n$ in $N$ are reachable
by a sequence of edges labelled by $e_1\dots e_k$ and connecting states
$\ii n,n_1,\dots,n_{k-1},n$, such that
\[
n_1 = t(\ii n,e_1), \quad n_i = t(n_{i-1},e_i),\ i = 2,\dots,k-1,\quad
n= t(n_{k-1},e_k).
\]
%
A finite sequence of events $s\in\Sigma^*$ is a \emph{trace} of $P$, if and only if
there
is a path through $G(P)$ starting  at $\ii n$ whose edge labels coincide with
$s$. The set of traces of $P$ is denoted by $\trc(P)$. We use the analgous notation
$\trc(G(P))$ for the set of finite, initialised paths through $G(P)$, each path represented by its finite sequence of edge labels.
If $s\in\trc(P)$, then
the process corresponding to $P$ after having executed $s$ is denoted by
$P/s$. Since $G(P)$ is normalised, there is a unique node reached by applying
the events from $s$ one by one, starting in $\ii n$. Therefore, $G(P)/s$  is
also well defined. By $[n]^0$ we denote the \emph{initials} of $n$:~the set
of events occurring as labels in any outgoing transitions.
$$
[n]^0 = \{ e\in\Sigma~|~(n,e)\in\dom~t \}
$$
We also use this notation for CSP processes:~$[P]^0$ is the set of events $P$
may engage into, in other words, the initials of $P$ after the empty trace of
events, that is, $initials(P/\langle\rangle)$ as defined
in~\cite{Roscoe2010}, for example.

@@

When investigating  tests for failures refinement, the notion of
\emph{acceptances}~\cite{Hennessy:1988:ATP:50497}, which is dual to refusals,
is also useful. An acceptance set of $P/s$ is a subset of the initials
$[P/s]^0$, i.e., a subset of events labelling  outgoing transitions of
$G(P)/s$. If the behaviour of  $P/s$ is deterministic, its only acceptance
equals $[P/s]^0$, because $P/s$ never refuses any of the events contained in
this set. If $P/s$ is nondeterministic, it internally chooses one of its
\emph{minimal acceptance sets} $A$ and never refuses any event in $A$, while
refusing the events in $\Sigma-A$. The acceptances of $P/s$ are denoted by
$\accs(P/s)$, and the minimal acceptances by $\minaccs(P/s)$. They satisfy
the following properties.
%
\begin{eqnarray}
A\in \minaccs(P/s) & \Leftrightarrow & \exists R\in\maxrefs(P/s) \wedge A = \Sigma-R
\label{eq:accref}
\\
\bigcup \{ A~|~A\in \accs(P/s)\} & = & [P/s]^0
\label{eq:accinitials}
\\
 X\in\accs(P/s) &\Leftrightarrow & A\in \minaccs(P/s) \wedge A\subseteq X \subseteq [P/s]^0
 \label{eq:accsubset}
\end{eqnarray}
%
%Note that (\ref{eq:accref}) can be regarded as a definition of minimal
%acceptances by means of maximal refusals. In this case,
%(\ref{eq:accinitials}) and (\ref{eq:accsubset}) are consequences of this
%definition and the fact that refusals are subset-closed. \fixme{alcc: are you
%sure? (4) does not talk about Acc. Is this important?}
%\fixme{jp: I'm sure and I can prove this. However, we don't need this fact, so I suggest to
%removed it.}
%
Exploiting formulas (\ref{eq:accref}), (\ref{eq:accinitials}), and
(\ref{eq:accsubset}), every node of a normalised transition graph can
alternatively be labelled with their minimal acceptances, and this
information is equivalent to that contained in the maximal refusals. Since
process states $P/s$ are equivalently expressed by states $G(P)/s$ of $P$'s
normalised transition graph, we also write $\minaccs(G(P)/s)$ and
$\maxrefs(G(P)/s)$ to denote the minimal acceptences and maximal refusals,
labelling graph state $G(P)/s$ and note that $\minaccs(G(P)/s) =
\minaccs(P/s)$ and $\maxrefs(G(P)/s) = \maxrefs(P/s)$.

@@

Refinement relations between finite-state CSP processes $P, Q$ can be be
expressed by means of their normalised transition graphs in the following
way. %\fixme{alcc: trc was defined for P, not G(P) and this lemma is used
%later for trc(P). So, we need to be clearer as to what we mean.}
%\fxnote{jp: I have now introduced notation $\trc(G(P))$ above, so I think that the
%statement of the lemma is now sufficiently clear.}
%
\begin{lemma}
  \label{lemma:tgtrcref}
  \begin{eqnarray}
  P \lessdet_T Q & \Leftrightarrow & \trc(G(Q)) \subseteq\trc(G(P))
  \label{eq:trcrefa}
  \\
  \label{eq:failrefa}
  P \lessdet_F Q & \Leftrightarrow & \trc(G(Q)) \subseteq\trc(G(P)) \wedge {} \nonumber
  \\ & & \forall s\in\trc(G(Q)), R_Q\in\maxrefs(G(Q)/s):  \nonumber
  \\ & & \tabd
  \exists R_P\in\maxrefs(G(P)/s): R_Q\subseteq R_P
  \\
  \label{eq:failrefb}
  P \lessdet_F Q & \Leftrightarrow & \trc(G(Q)) \subseteq\trc(G(P)) \wedge {} \nonumber
  \\ & & \forall s\in\trc(G(Q)), A_Q\in\minaccs(G(Q)/s): \nonumber
  \\ & & \tabd
  \exists A_P\in\minaccs(G(P)/s): A_P\subseteq A_Q
  \end{eqnarray}
  \xbox
\end{lemma}
\fixme{alcc: I'll write an explanation of these results.}

Given a finite-state CSP process $P$ and its normalised transition graph
\[
G(P) = ( N, \ii n, \Sigma, t : N\times\Sigma \pfun N, r : N \fun \mathbb{P}\mathbb{P}(\Sigma)),
\]
suppose that $V\subseteq\Sigma^*$ is a
prefix-closed set  of sequences of events. By $t(\ii n,V)$ we denote the set
\[
t(\ii n,V) = \{ n\in N~|~\exists s\in V: s\in\trc(P)\wedge G(P)/s = n \}
\]
of nodes in $N$ that are reachable in $G(P)$ by applying traces of $V$. The
lemma below specifies a construction method for such sets $V$ reaching {\it
every} node of $N$.

\begin{lemma}
\label{lemma:extendV} Let $P$ be a CSP process with normalised transition
graph $G(P)$. %\fixme{alcc: can a
%normalised graph for a process have states that are unreachable?}
%\fxnote{jp: You are rights: all states in G(P) are reachable by construction of the graph -- I have added this information to the section introducing the graphs.}
Let
$V\subseteq\Sigma^*$ be a finite prefix-closed set of sequences of events.
Suppose that  $G(P)$ reaches $k < |N|$ nodes under $V$, that is, $|t(\ii
n,V)| = k$. Let $V.\Sigma$ denote the set of all sequences from $V$, extended
by any event of $\Sigma$. Then $G(P)$ reaches at least $(k+1)$ nodes under
$V\cup V.\Sigma$.
\end{lemma}
\begin{proof}
Suppose that $n'\in (N - t(\ii n,V))$.  Since all nodes in $N$ are reachable,
there exists a trace $s$ such that $G(P)/s = n'$. Decompose $s = s_1.e.s_2$
with $s_i\in\Sigma^*, e\in\Sigma$, such that $G(P)/s_1 \in t(\ii n,V)$ and
$G(P)/s_1.e \not\in t(\ii n,V)$. Such a decomposition always exists, because
$V$ is prefix-closed and therefore contains the empty trace $\varepsilon$.
Note, however, that it is not necessarily the case that $s_1\in V$.

Since $G(P)$ reaches $G(P)/s_1$ under $V$, there exists a trace $u\in V$ such
that $G(P)/u = G(P)/s_1 = \ol n$. Since $s = s_1.e.s_2$ is a trace of $P$ and
$G(P)/s_1 = \ol n$, then $(\ol n,e)$ is in the domain of $t$. So, $ G(P)/u.e
= G(P)/s_1.e = n$ is a well-defined node of $N$ not contained in $t(\ii
n,V)$. Since $u.e\in V\cup V.\Sigma$, $G(P)$ reaches at least the additional
node $n$ under $V\cup V.\Sigma$. This completes the proof. \xbox
\end{proof}
%\fixme{alcc: Explain the structure of the section? Can the above result be in
%a section, rather than the introduction of the section?}
%\fxnote{jp: I have moved this lemma now to the section on preliminaries, this should make the main section on the finite complete test suite easier to read.}

@@

To prove ``$2 \Rightarrow 1$'', assume $P\lessdet_T Q$, but $P\not\lessdet_F
Q$. According to Lemma~\ref{lemma:tgtrcref}, (\ref{eq:failrefb}), there
exists $s\in\trc(P)\cap \trc(Q)$ such that
\[
\exists A_Q\in\minaccs(G(Q)/s): \forall A_P\in\minaccs(G(P)/s): A_P\not\subseteq A_Q
\qquad (*)
\]
Let $A$ be such a set $A_Q$ fulfilling (*).
Define
\[
\overline H = \bigcup\{ A_P \setminus A~|~A_P\in\minaccs(G(P)/s) \}.
\]
Since $A_P \setminus A \neq\varnothing$ for all $A_P$ because of (*),
$\overline H$ is a hitting set of $\minaccs(G(P)/s)$ which has an  empty
intersection with $A_Q$. %\fixme{alcc: minaccs is used for P above.}
%\fxnote{jp: Above Example 1, I have now introduced the notation $\minaccs(G(P)/s)$ and
%$\maxrefs(G(P)/s)$.}
Minimising $\overline H$ induces the existence of a minimal hitting set $H\in
\text{minHit}(P/s)$ which is {\it not} a hitting set of $\minaccs(G(Q)/s)$, a
contradiction to Assumption~2. This completes the proof of the lemma. \xbox
\end{proof}
%
This result is used in the sequel in Section~\ref{sec:finitecompletefails}.
